A Fully Decentralized Infrastructure for Subscription-based IoT Data Trading

### 2020-08-11 paper.tex

\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{float}
\usepackage{tabularx}
\usepackage{makecell}
\usepackage{multirow}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic,algorithm}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{listings}
\graphicspath{{./images/}}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

@@

\section{Introduction}
In IoT, the development of M2M technology\cite{M2M}, cyber physical system (CPS)\cite{CPS}, and Industry 4.0 grow rapidly. As the physical and digital data are deeply intertwined, the interactions among digital twins act as data exchanges\cite{digitaltwin} which brings potential value to IoT applications, such as health care\cite{healthCare}, factories, and vehicles\cite{AutonomousDriving}, and brings up new business models where data is considered as tradeable digital assets. With the diverse data streams generated by different entities and carried across organizations among expanding the number of interconnected devices, it is a challenge for data holders to share and track their data assets. Therefore, data trading platform is viewed as a solution to build a secure, reliable, and scalable data sharing mechanism where data providers and data consumers meet.

@@

% SaaS
The economic incentive model in our proposed IoT data trading platform works as a subscription service that valuates the payments by data usage. The subscription-based models (e.g. SaaS and platform-as-a-service (PaaS)) are embedded in our lives, such as newspaper, video/audio streaming, and software, which can bring enormous operating incomes and offer advantages that allows service providers have the flexibility in resource planning as well as a better prediction of revenue streams. And for service consumers, they can subscribe and unsubscribe the services at will at any time. The subscription-based model can easily adapt the underlying pub/sub communication model and benefit both data providers and consumers which promotes good circulation of the platform. In \cite{SaaS}, the author pointed out that the strength of the relationships between consumers and providers of subscription-based services depends on the service quality (i.e., the data quality in data trading platform), and trust is key to successful subscription-based services.

%trustless & automated
%trust: 對系統、對身份、對資料
In data trading platforms, there are three major aspects that need trust: identity, data ownership, and trading. Proving the identity as well as the ownership of data is a challenging issue among various roles in a large scale platform. In current systems, a trusted third-party authority like certificate authority (CA) is widely adopted to certify each member. However, building trust upon CA is dangerous and fragile. The compromise of routers or CA can break the trust of the entire system, and the data portability of using third-party services is still under suspicion. Meanwhile, these concerns are taking into data storage as well, keeping data assets in a centralized data storage or cloud service meet the constraints of data portability, which may violate the General Data Protection Regulation (GDPR)\cite{GDPR}. Besides CA and storage, the subscription history of data consumers may reveal his/her interests and the exposure of personal information during trading process against GDPR as well. The final trust is trading procedures, apart from setting up the agreement between providers and consumers, both of them need to confirm that the agreement will be executed even if one party fails to comply.

@@

\subsection{Contributions}
The contributions in the paper are summarized as follow:
\begin{itemize}
    \item We investigate the usage of DLTs and distributed data storages in order to propose a decentralized architecture for IoT trading and to enable participants to automate the trading procedures and secure digital assets exchange across different organizations.
    \item We apply the subscription-based services economic model to adapt the features of IoT streaming data and clarify the importance of using a decentralized identity system for trading.
    \item We evaluate the performance and concerns of adopting an end-to-end encrypted transmission protocol, and propose an alternative approach for low-level devices to adopt this protocol while ensuring the privacy of using the protocol.
    \item We analyze the cost of designed smart contract to estimate the minimum data product price for data providers to sustain a product and the minimum cost for data consumers to interact with data products.
\end{itemize}

\subsection{Paper Organization}
The rest of this paper is organized as follows. In Section~\ref{section:relatedWork}, some related work of pub/sub over blockchain, decentralized data trading infrastructures, and distributed data storage are analyzed. In Section~\ref{section:design_thinking}, the components of proposed subscription-based data trading architecture is illustrated. In Section~\ref{section:MAM}, the end-to-end encrypted transmission protocol is introduced along with the protocol delegation solution for low-level devices. In Section~\ref{section:trading_model}, the trading models of proposed architecture are well explained. The analysis results are given in Section~\ref{section:evaluation}. Finally, we conclude the paper in Section~\ref{section:conclusion} and the future work in Section~\ref{section:future_work}.

@@

Filecoin \cite{FileCoin} in the Inter-Planetary File system (IPFS) is an incentive layer to incent nodes to provide storage. IPFS is a content-based addressing storage model in a peer-to-peer network, in which users can obtain the data with the unique hash value through the network. Furthermore, considering the data may lose in the network, IPFS provides \textbf{pin} command for users to store the contents locally that will not be removed. However, the IoT sensor devices have limited space which is not appropriate to preserve generated data. Also, users can have a hard time finding the contents quickly if data is not widely available, the large amount of IPFS nodes and contents. Thus, several pinning services like Pinata provide multiple IPFS nodes around the world to pin users' contents and charge by the size and replicas of contents, and users can even join the dedicated network to speed up the retrieving process. Nevertheless, the efficiency problems still occur if the dedicated network grows larger.

@@

\subsection{Unsubscribe to Data Products}
Data consumers can unsubscribe to data products when they're not interested anymore. Product Contract marks the consumer as not subscribing, corresponding to the parameter $isSubscribe$ in $Product$ structure since the delete function in Solidity does not remove the actual element but resetting it to default value. When data consumers trigger the unsubscribe event, a withdraw function is executed to pay a subscription fee to providers and return the rest of the fee back to consumers. The amount of payment distributed to different players is defined in formula~\ref{equation:unsubscribe_provider}, \ref{equation:unsubscribe_broker}, \ref{equation:unsubscribe_consumer}. $i$ is the number of data when refunding condition is met, $price$  is the subscription price, $M$ is the number of expected data samples, $F_{b}$ (\%) is the brokerage fee which is expressed as a percentage, $F_{t}$ is the transaction fee of the smart contract, and $F_{cancel}$ is the cancellation fee charged for unsubscription.

@@

Note that data uploading is only related to MAM operations on IOTA, which does not require any fees. Therefore, the Gas estimation only includes trading processes. Creating a Product Contract (Listing \ref{lst:constructor}) costs the most Gas (0.164 - 0.596 Ether), providers should have enough Ethers to launch the product at the start. Functions (Listing \ref{lst:key_exchange} (0.012 - 0.044 Ether) and Listing \ref{lst:key_certification}) (0.022 - 0.0834 Ether) that are related to key certification and key exchange need a lot of gas as well. Storage on blockchain is expensive. For instance, an SSTORE operation may cost 20000 gas when the value is set to non-zero from zero according to Ethereum Yellow Paper\cite{Ethereum}. However, in our trading model, we preserve a lot of information in the Product Contract in order to record as more details as possible during trading procedures, such as ciphertexts,  signatures during key certification and key exchange. Especially, in the key exchange protocol, it takes several steps of communications between brokers, data providers, and data consumers via storing states to Product Contract. Therefore, it is worth to explore the usage of the non-interactive zero-knowledge protocol such as zk-SNARKS(zero-knowledge succinct non-interactive argument of knowledge)\cite{Snark} which can generate proofs off-chain and simplify the verification process that could not only reduce network and storage cost but also improve the privacy and security of encryption keys. However, using the zero-knowledge protocol is out of the scope of this paper.

Base on Gas estimation in Table.~\ref{tab:gas}, we calculate the exact Ether with Formula~\ref{equa:ether_calculation} that data providers and data consumers need during the trade in Table.~\ref{tab:ether}. We combine the Gas usage of providers and brokers since the brokerage fee is also a critical cost to sustain the trade, which providers should considered. We conclude the four situations during the trading process: w/o refund and unsubscribe, w/ refund only, w/unsubscribe, and w/ refund and unsubscribe. The condition of w/o refund and unsubscribe represents the basic cost of  triggering regular functions, such as launch, add encryption key.

@@

When a consumer request a refund, he/she need to pay 0.0003 Ether additionally. Then if the refund established, all participants will withdraw the payment, which is already included in the basic cost. Observing the situation of w/ unsubscribe and w/ refund and unsubscribe, it does not cost them much more Ether for consumers. However, the data provider needs to update all encryption keys for remaining consumers that \textit{update encrypt key} will be triggered $N$ times, where $N$ is the number of remaining subscribers. This would cost data providers a lot if the unsubscription is frequent or many consumers subscribe to the data product. Hence, an efficient, low-cost, and secure way to perform access control while recording the unsubscription process transparent is important. An alternative strategy coming up with the latest development of MAM, which is renaming to Streams\cite{stream}, provides new functionality for subscribers to unsubscribe to data streams, thus, it is possible to perform access control directly through it rather than uploading new encryption key to the Product Contract. Nevertheless, Streams is still in an alpha version, the details and protocol still need to be well studied.

\begin{table*}[h]
    \caption{Data Price Estimation}
    \label{tab:ether}
    \centering
    \resizebox{0.8\textwidth}{!}{
    \begin{tabular}{|c|c|c|c|c|c|}
        \hline
        \textbf{condition} & \textbf{sender} & \textbf{\makecell{minimum price \\ (Ether)}} & \textbf{\makecell{minimum price \\ (USD)}} & \textbf{\makecell{maximum price \\ (Ether)}} & \textbf{\makecell{maximum price \\ (USD)}} \\
        \hline
        \multirow{2}{*}{\makecell{w/o refund \\ unsubscribe}} & provider & 0.021 & 5.038 & 0.076 & 18.291 \\
         & consumer & 0.00061 & 0.145 & 0.0022 & 0.527\\
        \hline
        \multirow{2}{*}{w/ refund} & provider & 0.021 & 5.038 & 0.076 & 18.291 \\
         & consumer & 0.00069 & 0.166 & 0.0025 & 0.603 \\
        \hline
        \multirow{2}{*}{w/ unsubscribe} & provider & 0.021 + 0.002N & 5.038 + 0.497N & 0.076  + 0.007N & 18.291 + 1.806N \\
         & consumer & 0.00067 & 0.159 & 0.0024 & 0.579 \\
        \hline
        \multirow{2}{*}{\makecell{w/ refund \\ unsubscribe}} & provider & 0.021 + 0.002N & 5.038 + 0.497N & 0.076  + 0.007N & 18.291 + 1.806N\\
         & consumer & 0.0007 & 0.180 & 0.0027 & 0.655 \\
        \hline
    \end{tabular}
    }
\end{table*}

>>>>

\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{float}
\usepackage{tabularx}
\usepackage{makecell}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic,algorithm}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{listings}
\graphicspath{{./images/}}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

@@

\section{Introduction}
In IoT, the development of M2M technology\cite{M2M}, cyber physical system (CPS)\cite{CPS}, and Industry 4.0 grow rapidly. As the physical and digital data are deeply intertwined, the interactions among digital twins act as data exchanges\cite{digitaltwin} which brings potential value to IoT applications, such as health care\cite{healthCare}, factories, and vehicles\cite{AutonomousDriving}, and brings up new business models where data is considered as tradeable digital assets. With the diverse data streams generated by different entities and carried across organizations among expanding number of interconnected devices, it is a challenge for data holders to share and track their data assets. Therefore, data trading platform is viewed as a solution to build a secure, reliable, and scalable data sharing mechanism where data providers and data consumers meet.

@@

% SaaS
The economic incentive model in our proposed IoT data trading platform works as a subscription service that valuates the payments by data usage. The subscription-based models (e.g. SaaS and platform-as-a-service (PaaS)) are embedded in our lives, such as newspaper, video/audio streaming, and software, which can bring enormous operating incomes and offer advantages that allows service providers to have the flexibility in resource planning as well as a better prediction of revenue streams. And for service consumers, they can subscribe and unsubscribe the services at will at any time. The subscription-based model can easily adapt the underlying pub/sub communication model and benefit both data providers and consumers which promotes good circulation of the platform. In \cite{SaaS}, the author pointed out that the strength of the relationships between consumers and providers of subscription-based services depends on the service quality (i.e., the data quality in data trading platform), and trust is key to successful subscription-based services.

%trustless & automated
%trust: 對系統、對身份、對資料
In data trading platforms, there are three major aspects that need trust: identity, data ownership, and trading. Proving the identity as well as the ownership of data is a challenging issue among various roles in a large scale platform. In current systems, a trusted third-party authority like certificate authority (CA) is widely adopted to certify each member. However, building trust upon CA is dangerous and fragile. The compromise of routers or CA can break the trust of the entire system, and the data portability of using third-party services is still under suspicion. Meanwhile, these concerns are taking into data storage as well, keeping data assets in a centralized data storage or a cloud service meet the constraints of data portability, which may violate the General Data Protection Regulation (GDPR)\cite{GDPR}. Besides CA and storage, the subscription history of data consumers may reveal his/her interests and the exposure of personal information during trading process against GDPR as well. The final trust is trading procedures, apart from setting up the agreement between providers and consumers, both of them need to confirm that the agreement will be executed even if one party fails to comply.

@@

\subsection{Contributions}
The contributions in the thesis are summarized as follow:
\begin{itemize}
    \item We investigate the usage of DLTs and distributed data storages in order to propose a decentralized architecture for IoT trading and to enable participants to automate the trading procedures and secure digital assets exchange across different organizations.
    \item We apply the subscription-based services economic model to adapt the features of IoT streaming data and clarify the importance of using a decentralized identity system for trading.
    \item We evaluate the performance and concerns of adopting an end-to-end encrypted transmission protocol, and propose an alternative approach for low-level devices to adopt this protocol while ensuring the privacy.
    \item We analyze the cost of the designed smart contract to estimate the minimum data product price for data providers to sustain a product and the minimum cost for data consumers to interact with data products.
\end{itemize}

\subsection{Thesis Organization}
The rest of this thesis is organized as follows. In Section~\ref{section:relatedWork}, some related work of pub/sub over blockchain, decentralized data trading infrastructures, and distributed data storage are analyzed. In Section~\ref{section:design_thinking}, the components of proposed subscription-based data trading architecture is illustrated. In Section~\ref{section:MAM}, the end-to-end encrypted transmission protocol is introduced along with the protocol delegation solution for low-level devices. In Section~\ref{section:trading_model}, the trading models of proposed architecture are well explained. The analysis results are given in Section~\ref{section:evaluation}. Finally, we conclude the thesis in Section~\ref{section:conclusion} and the future work in Section~\ref{section:future_work}.

@@

Filecoin \cite{FileCoin} in the Inter-Planetary File system (IPFS) is an incentive layer to incent nodes to provide storage. IPFS is a content-based addressing storage model in a peer-to-peer network, in which users can obtain the data with the unique hash value through the network. Furthermore, considering the data may lose in the network, IPFS provides \textbf{pin} command for users to store the contents locally that will not be removed. However, the IoT sensor devices have limited space which is not appropriate to preserve generated data. Also, users can have a hard time finding the contents quickly if data is not widely available, and the large amount of IPFS nodes and contents. Thus, several pinning services like Pinata provide multiple IPFS nodes around the world to pin users' contents and charge by the size and replicas of contents, and users can even join the dedicated network to speed up the retrieving process. Nevertheless, the efficiency problems still occur if the dedicated network grows larger.

@@

\subsection{Unsubscribe to Data Products}
Data consumers can unsubscribe to data products when they're not interested anymore. Product Contract marks the consumer as not subscribing, corresponding to the parameter $isSubscribe$ in $Product$ structure since the delete function in Solidity does not remove the actual element but resetting it to default value. When data consumers trigger the unsubscribe event, a withdraw function is executed to pay a subscription fee to providers and return the rest of the fee back to consumers. The amount of payment distributed to different players is defined in formula~\ref{equation:unsubscribe_provider}, \ref{equation:unsubscribe_broker}, \ref{equation:unsubscribe_consumer}. $i$ is the number of data when unsubscription request is launched, $price$  is the subscription price, $M$ is the number of expected data samples, $F_{b}$ (\%) is the brokerage fee which is expressed as a percentage, $F_{t}$ is the transaction fee of the smart contract, and $F_{cancel}$ is the cancellation fee charged for unsubscription.

@@

Note that data uploading is only related to MAM operations on IOTA, which does not require any fees. Therefore, the Gas estimation only includes trading processes. Creating a Product Contract (Listing \ref{lst:constructor}) costs the most Gas (0.164 - 0.596 Ether), providers should have enough Ethers to launch the product at the start. Functions (Listing \ref{lst:key_exchange} (0.012 - 0.044 Ether) and Listing \ref{lst:key_certification} (0.022 - 0.0834 Ether)) that are related to key certification and key exchange need a lot of gas as well. Storage on blockchain is expensive. For instance, an SSTORE operation may cost 20000 gas when the value is set to non-zero from zero according to Ethereum Yellow Paper\cite{Ethereum}. However, in our trading model, we preserve a lot of information in the Product Contract in order to record as more details as possible during trading procedures, such as ciphertexts,  signatures during key certification and key exchange. Especially, in the key exchange protocol, it takes several steps of communications between brokers, data providers, and data consumers via storing states to Product Contract. Therefore, it is worth to explore the usage of the non-interactive zero-knowledge protocol such as zk-SNARKS(zero-knowledge succinct non-interactive argument of knowledge)\cite{Snark} which can generate proofs off-chain and simplify the verification process that could not only reduce network and storage cost but also improve the privacy and security of encryption keys. However, using the zero-knowledge protocol is out of the scope of this paper.

Base on Gas estimation in Table.~\ref{tab:gas}, we calculate the exact Ether with Formula~\ref{equa:ether_calculation} that data providers and data consumers need during the trade in Table.~\ref{tab:ether}. We combine the Gas usage of providers and brokers since the brokerage fee is also a critical cost to sustain the trade, which providers should considered. We conclude the three situations during the trading process: 1.) basic cost of a trade, including product launching, subscribing, encryption key uploading and withdraw. 2.) refund operation, the fees needed for launching refund. 3.) unsubscribe operation, the fees needed for launching unsubscription.

@@

When a consumer request a refund, he/she need to pay 0.0008 - 0.0003 Ether additionally. Then if the refund established, all participants will withdraw the payment, which is already included in the basic cost. Observing the situation of unsubscribe operation, it does not cost much Ether for consumers. However, the data provider needs to update all encryption keys for remaining consumers that \textit{update encrypt key} will be triggered $N$ times, where $N$ is the number of remaining subscribers. This would cost data providers a lot if the unsubscription is frequent or many remaining consumers are left. Hence, an efficient, low-cost, and secure way to perform access control while recording the unsubscription process transparent is important. An alternative strategy coming up with the latest development of MAM, which is renaming to Streams\cite{stream}, provides new functionality for subscribers to unsubscribe to data streams, thus, it is possible to perform access control directly through it rather than uploading new encryption key to the Product Contract. Nevertheless, Streams is still in an alpha version, the details and protocol still need to be well studied.

\begin{table*}[h]
    \caption{Data Price Estimation}
    \label{tab:ether}
    \centering
    \resizebox{0.8\textwidth}{!}{
    \begin{tabular}{|c|c|c|c|c|c|}
        \hline
        \textbf{condition} & \textbf{sender} & \textbf{\makecell{minimum price \\ (Ether)}} & \textbf{\makecell{minimum price \\ (USD)}} & \textbf{\makecell{maximum price \\ (Ether)}} & \textbf{\makecell{maximum price \\ (USD)}} \\
        \hline
        \multirow{2}{*}{basic cost} & provider & 0.021 & 5.038 & 0.076 & 18.291 \\
         & consumer & 0.00061 & 0.145 & 0.0022 & 0.527\\
        \hhline{|=|=|=|=|=|=|}
        \multirow{2}{*}{refund} & provider & 0 & 0 & 0 & 0 \\
         & consumer & 0.00008 & 0.0209 & 0.0003 & 0.0761 \\
        \hline
        \multirow{2}{*}{unsubscribe} & provider & 0.002N & 0.497N & 0.007N & 1.806N \\
         & consumer & 0.00006 & 0.0143 & 0.0002 & 0.0519 \\
        \hline
    \end{tabular}
    }
\end{table*}
